---
title: "p8105_hw6_lg3158"
author: "Ling"
date: "11/27/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)
```

## Problem 1

- Data cleaning
  - transform categorical variables as factor (`babysex`,`frace`,`malform`,`mrace`)
  - check for missing values

```{r}
p1 = read_csv("https://www.p8105.com/data/birthweight.csv",
              col_names = TRUE) %>%
  mutate(babysex = recode_factor(as_factor(babysex), "1" = "male","2" = "female"),
         frace = recode_factor(as_factor(frace), "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
         malform = recode_factor(as_factor(malform), "1" = "present", "0" = "absent"),
         mrace = recode_factor(as_factor(mrace), "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"))
anyNA(p1)
```

- Data description
  - there are `r ncol(p1)` variables and `r nrow(p1)` observations in the data set, without any missing values. 
  
- Regression model
  - by common sense, gestational age in weeks, baby's head circumference and length at birth are associated with birth weight, though these are not clinically meaningful.
  - thus, drop `bhead`, `blength`, and `gaweeks` and use the remained variables to construct the linear regression model using step-wise selection (forward process).
  - rerun the model using `lm`.
  - produce the plot of model residuals against fitted values.
```{r}
p1_reg = p1 %>%
         select(-c(bhead, blength, gaweeks))

intercept_only_p1 = lm(bwt ~ 1, data = p1_reg)
full_p1 = lm(bwt ~ ., data = p1_reg)
forward_p1 = step(intercept_only_p1, direction = "forward", scope = formula(full_p1), trace = 0)

foward_anova = as_tibble(forward_p1$anova)
p1_reg_summary = p1_reg %>%
  lm(bwt ~ mrace + delwt + smoken + ppbmi + babysex + ppwt + mheight + fincome, data = .)

p1_reg %>%
  modelr::add_residuals(p1_reg_summary) %>%
  modelr::add_predictions(p1_reg_summary) %>%
  ggplot(aes(x = pred, y = resid)) +
    geom_point() +
    theme_minimal()
```

- Cross validation

```{r}
cv_df = crossv_mc(p1 , 100) %>%
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble),
         p1_linear = map(train, ~lm(bwt ~ mrace + delwt + smoken + ppbmi + babysex + ppwt + mheight + fincome, data = .)),
         p1_2 = map(train, ~ lm (bwt ~ blength + gaweeks, data = .)),
         p1_3 = map(train, ~ lm (bwt ~ bhead + blength + babysex + bhead*babysex + blength*babysex + bhead*blength, data = .)),
         rmse_linear = map2_dbl(p1_linear, test, ~ rmse(model = .x, data = .y)),
         rmse_2 = map2_dbl(p1_2, test, ~ rmse(model = .x, data = .y)),
         rmse_3 = map2_dbl(p1_3, test, ~ rmse(model = .x, data = .y))) %>%
  select(starts_with('rmse')) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse, fill = model, color = model)) + 
    geom_violin() +
    theme_minimal()

cv_df
```

- From the violin plot we can see that the model using head circumference, length, sex and all the interactions has the minimal rmse, indicating that this model fits the data best among all 3 models.

## problem 2

- Steps:
  - import the data using the given code chunk
  - bootstrap and data wrangling the results
    - the summary data for bootstrap is named as `bootstrap`
    - create 5000 bootstrap samples
    - use `map` to create the estimates
      - log(beta_0 * beta_1) is in `results` list
      - r^2 is in `rsquares` list
    - data wrangling
      - unnest the `results` list, keep the variable `term` and `estimate`
      - transform the `intercept` and `tmin` into wider format
      - calculate log(beta_0 * beta_1), the result variable is named as `log_betas`
      - unnest the `rsquare` list, extract `r_squared` variable
  - density plot (named as `bootstrap_density`)
    - transform the `bootstrap` data into long form, in which:
      - `term`: whether the variable is `log_betas` or `r_squared`
      - `estimate`: the estimated values
    - create the density plot, group by `term`
    
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

bootstrap = weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, ~ lm (tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    rsquares = map(models, broom::glance)
    ) %>%
  select(results, rsquares) %>%
  unnest(results) %>%
  select(term, estimate, rsquares) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  unnest(rsquares) %>%
  janitor::clean_names() %>%
  select(intercept, tmin, r_squared) %>%
  mutate(log_betas = log(intercept * tmin))

bootstrap_density = bootstrap %>%
  select(-intercept, -tmin) %>%
  pivot_longer(cols = c(log_betas, r_squared), names_to = "term", values_to = "estimate") %>%
  ggplot(aes(x = estimate, group = term, color = term)) +
    geom_density() +
    theme_minimal() +
    scale_color_discrete(labels = c("log(beta_0 * beta_1)", "r^2"))

bootstrap_density
``` 

  - Description of the density plot
    - the 2 values are approximately normally distributed, with different means
    - the mean of r^2 is larger than the mean of log(beta_0 * beta_1) (2 vs. 0.9)
    - the range of log(beta_0 * beta_1) is narrower than r^2

- calculate 95% CI (named as `bootstrap_summary`)
  - transform the `bootstrap` data into long form, in which:
    - `term`: whether the variable is `log_betas` or `r_squared`
    - `estimate`: the estimated values
    - create the density plot, group by `term`

```{r}
bootstrap_summary = bootstrap %>%
  select(-intercept, -tmin) %>%
  pivot_longer(cols = c(log_betas, r_squared), names_to = "term", values_to = "estimate") %>%
  group_by(term) %>%
  summarize(ci_lower = quantile(estimate, 0.025),
          ci_upper = quantile(estimate, 0.975))

knitr::kable(bootstrap_summary, digits = 3)
```




